---
title: "Omics Playground"
output: 
  flexdashboard::flex_dashboard:
    theme: default
    orientation: col
    logo: resources/irb-small.png
    favicon: resources/irb-small.png
##logo: resources/bigomics-logo-white-48px.png
##  favicon: resources/bigomics-logo-white-48px.png
    css: resources/custom.css
runtime: shiny
---


```{r global=TRUE, echo=FALSE, message=FALSE, include=FALSE}
###```{r global, include=FALSE} 
## load data in 'global' chunk so it can be shared by all users of the dashboard
library(shiny)
library(rmarkdown)

## set global folders
all.dir = c("R","lib","pgx")
ROOT=NULL
if(all(dir.exists(paste0("/playground",all.dir)))==TRUE) ROOT="/playground"
if(all(dir.exists(paste0("../",all.dir)))==TRUE) ROOT=".."
if(all(dir.exists(all.dir))==TRUE) ROOT="."
ROOT
if(is.null(ROOT)) stop("FATAL ERROR: could not fine R/lib/pgx directories")
cat("setting ROOT to",ROOT,"\n")

RDIR = paste0(ROOT,"/R")
FILES = paste0(ROOT,"/lib/")
PGX.DIR = paste0(ROOT,"/pgx/")
cat("setting RDIR =",RDIR,"\n")
cat("setting FILES =",FILES,"\n")
cat("setting PGX.DIR =",PGX.DIR,"\n")

source("./pgx-init.R", local=TRUE)

if(0) {
    load("../pgx/GSE10846-dlbcl.pgx")    
    ngs <- pgx.initialize(ngs)    
}
```

```{r child = 'module-intro.Rmd'}
```

```{r child = 'module-datatable.Rmd'}
```

```{r child = 'module-clustering.Rmd'}
```

```{r child = 'module-expression.Rmd'}
```

```{r child = 'module-enrichment.Rmd'}
```

```{r child = 'module-functional.Rmd'}
```

```{r child = 'module-intersect.Rmd'}
```

```{r child = 'module-signature.Rmd'}
```

```{r child = 'module-predict.Rmd'}
```

```{r child = 'module-profiling.Rmd'}
```

`r if(!PRO.VERSION) {"<!-------------------------------------------"}`

```{r child = 'module-batchcorrection.Rmd'}
```

`r if(!PRO.VERSION) {"--------------------------------------------->"}`


`r if(!DEV.VERSION) {"<!-------------------------------------------"}`

```{r child = 'module-system.Rmd'}
```

`r if(!DEV.VERSION) {"--------------------------------------------->"}`
