Biomarker
================================================================================

Inputs {.sidebar data-width=250}
--------------------------------------------------------------------------------

<br> **Biomarker Module.** Select biomarkers that can be used for
classification or prediction purposes. The phenotype of interest can
be multiple categories (classes) or patient survival data. 


```{r}
htmlOutput("curdata10", class="current-data")
##----------------------------------------------------------------------
## More Info (pop up window)
##----------------------------------------------------------------------
tipify( actionLink("pdx_info", "Info", icon = icon("info-circle")),
       "Show more information about this module.")

pdx_infotext =
    "The <strong>Biomarker Module</strong> performs the biomarker selection that can be used for classification or prediction purposes.

<br><br>To better understand which genes, mutations, or gene sets influence the final phenotype the most, Playground calculates a variable importance score for each feature using state-of-the-art machine learning algorithms, including LASSO, elastic nets, random forests, and extreme gradient boosting, and provides the top 50 features according to cumulative ranking by the algorithms. By combining several methods, the platform aims to select the best possible biomarkers.

<br><br>The phenotype of interest can be multi-categorical classes or patient survival data. Instead of choosing a phenotype, users can also specify a particular contrast from the analysis and perform biomarker selection. The platform also provides a heatmap of samples based on identified top features. 

<br><br>In addition, it generates a classification tree using top features and provides expression boxplots by phenotype classes for features present in the tree. The platform can also provide a survival tree analysis using top features and provides expression boxplots by phenotype classes for features present in the tree.

"

observeEvent( input$pdx_info, {
    dbg("<module-biomarker::observe pdxinfo> reacted")
    showModal(modalDialog(
        title = HTML("<strong>Biomarker Module</strong>"),
        HTML(pdx_infotext),
        easyClose = TRUE ))
})
```

<br><hr>

```{r}
tipify(selectInput("pdx_predicted","Predicted target:", choices=NULL),
       "Select the target variable for biomarker selection.",
       placement="top")

tipify( selectInput("pdx_level","Feature level:", choices=c("gene","geneset")),
       "Select feature level: gene or geneset", placement="top")

tipify( selectInput("pdx_filter","Feature filter:", choices=NULL),
       "Select a filter for the features.", placement="top")

conditionalPanel(
    "input.pdx_filter == '<custom>'",
    tipify(
        ##selectizeInput("pdx_select","Custom features:", choices=NULL, multiple=TRUE),
        textAreaInput("pdx_select", "Custom features:", value = NULL,
                      rows=8, placeholder="Paste your gene list"),
        "Paste a custom gene list to be used as features.", placement="top")
)

tipify(actionButton("pdx_runbutton", label="Run", class="blinking"),
       "Click to start biomarker computation.", placement="right")

br();hr()

if(DEV.VERSION) {
    conditionalPanel(
        "input.pdx_options % 2 == 1",
        tagList(
            h6("Developer options:"),
            checkboxInput('pdx_multiomics','multi-omics weighting',FALSE)
        )
    )
}

##input_pdx_select <- reactive({
##    input$pdx_select
##}) %>% debounce(3000)

input_pdx_select <- reactive({
    dbg("<input_pdx_select>  reacted")
    gg <- input$pdx_select
    if(is.null(gg)) return(NULL)
    gg <- strsplit(as.character(gg), split="[, \n\t]")[[1]]
    if(length(gg)==0) return(NULL)
    if(length(gg)==1 && gg[1]!="") gg <- c(gg,gg)  ## hack to allow single gene....
    return(gg)
}) %>% debounce(1000)


observe({
    ngs <- inputData()
    ##if(is.null(ngs)) return(NULL)
    req(ngs)
    dbg("<module-biomarker::observe1> reacted")

    ct <- colnames(ngs$Y)
    ct <- grep("group|sample|patient|donor",ct,value=TRUE,invert=TRUE)
    updateSelectInput(session, "pdx_predicted", choices=ct )
    dbg("<module-biomarker::observe1> done!")    
})

observe({
    ngs <- inputData()
    ##if(is.null(ngs)) return(NULL)
    req(ngs, input$pdx_level)
    dbg("<module-biomarker::observe2> reacted")

    if(input$pdx_level=="geneset") {
        ft <- names(COLLECTIONS)
        nn <- sapply(COLLECTIONS, function(x) sum(x %in% rownames(ngs$gsetX)))
        ft <- ft[nn >= 10]
    } else {
        ## gene level
        ft <- pgx.getFamilies(ngs,nmin=10,extended=FALSE)
    }
    ft <- sort(ft)
    if(input$pdx_level == "gene") ft = sort(c("<custom>",ft))
    updateSelectInput(session, "pdx_filter", choices=ft, selected="<all>")    

    dbg("<module-biomarker::observe2> done!")
})


```


Col {.tabset data-width=650}
----------------------------------------------------------------------------------------

### Importance
    
```{r message=FALSE, warning=FALSE, fig.width=4}

##calcVariableImportance <- reactive({
calcVariableImportance <- eventReactive( input$pdx_runbutton, {

    ## 
    ## This code also features a progress indicator.
    ##
    ##

    dbg("calcVariableImportance:: reacted on runbutton")
    
    ngs <- inputData()
    ##if(is.null(ngs)) return(NULL)
    req(ngs, input$pdx_predicted)
    
    ct=5
    ct=12
    colnames(ngs$Y)    
    ct <- input$pdx_predicted
    
    dbg("calcVariableImportance","predicting = ",ct,"\n")
    do.survival <- grepl("survival",ct,ignore.case=TRUE)
    dbg("calcVariableImportance","do.survival = ",do.survival,"\n")
        
    if(is.null(ct)) return(NULL)
    
    dbg("calcVariableImportance","2")

    NFEATURES=50
    NFEATURES=60

    # Create a Progress object
    progress <- shiny::Progress$new()
    # Make sure it closes when we exit this reactive, even if there's an error
    on.exit(progress$close())
    
    dbg("calcVariableImportance:: 1\n")
    progress$set(message = "Variable importance", value = 0)
    
    if(!(ct %in% colnames(ngs$Y))) return(NULL)
    y0 <- ngs$Y[,ct]
    names(y0) <- rownames(ngs$Y)
    y <- y0[!is.na(y0)]
    dbg("calcVariableImportance:: table.y=",table(y),"\n")
   
    table(y)
    if(length(y)<40) y <- head(rep(y,10),100)  ## augment to 100 samples
    table(y)

    dbg("calcVariableImportance:: 2\n")
    
    ##-------------------------------------------
    ## select features
    ##-------------------------------------------
    ## group prediction
    if(input$pdx_level=="geneset") {
        X <- ngs$gsetX[,names(y)]
    } else {
        X <- ngs$X[,names(y)]
    }

    dim(X)
    X0 <- X
    length(y)
    
    ## ----------- filter with selected features
    dbg("calcVariableImportance:: 3: filtering with selected features\n")
    progress$inc(1/10, detail = "Filtering features")
    
    ft="<all>"
    ft <- input$pdx_filter
    if(is.null(ft)) return(NULL)
    
    pp <- rownames(X)
    if(input$pdx_level=="geneset") {
        if(!(ft %in% names(COLLECTIONS))) return(NULL)
        pp <- COLLECTIONS[[ft]]
    } else {
        xfam <- c(names(ngs$families),names(GSETS))
        if(!ft %in% xfam) return(NULL)
        gg <- GSETS[[8]]
        gg <- NULL
        if(ft %in% names(ngs$families)) gg <- ngs$families[[ft]]
        if(ft %in% names(GSETS)) gg <- GSETS[[ft]]
        pp = filterProbes(ngs$genes, gg)
    }
    pp <- intersect(pp,rownames(X))
    X <- X[pp,,drop=FALSE]
    dim(X)
    
    ## ------------- filter with user selection
    ##sel <- input$pdx_select
    sel <- input_pdx_select()
    if(!is.null(sel) && length(sel)>0) {
        if(sel[1]!="") {
            pp <- intersect(rownames(X),sel)
            X <- X[pp,,drop=FALSE]
        }
    }

    dbg("calcVariableImportance:: 4a: dim.X=",dim(X),"\n")
        
    ## ----------- restrict to top 100
    dim(X)
    X <- head(X[order(-apply(X,1,sd)),,drop=FALSE], 10*NFEATURES)  ## top 100
    sdx <- mean(apply(X,1,sd))
    X <- X + 0.25*sdx*matrix(rnorm(length(X)),nrow(X),ncol(X))  ## add some noise
    dim(X)

    dbg("calcVariableImportance:: 4b: restrict top: dim.X=",dim(X),"\n")
    
    progress$inc(4/10, detail = "computing scores")
    
    ##-------------------------------------------
    ## compute importance values
    ##-------------------------------------------
    if(do.survival) {
        time <- abs(y)
        status <- (y > 0) ## dead is positive time
        methods=c("glmnet","randomforest","boruta","xgboost","pls")
        methods=c("glmnet","randomforest","xgboost","pls")
        P <- pgx.survivalVariableImportance(
            X, time=time, status=status, methods=methods)    
    } else {
        methods=c("glmnet","randomforest","boruta","xgboost","pls")
        methods=c("glmnet","randomforest","xgboost","pls")
        X1=X;y1=y
        names(y1) = colnames(X1) = paste0("x",1:ncol(X))
        P <- pgx.multiclassVariableImportance(X1, y1, methods=methods)    
        ##P <- pgx.variableImportance(X1, y1, methods=methods)
    }
    P <- abs(P)
    head(P)
    
    P[is.na(P)] <- 0
    P[is.nan(P)] <- 0
    P <- t( t(P) / (1e-3+apply(P,2,max,na.rm=TRUE)))
    ##P <- pmax(P,0.1)
    P <- P[order(-rowSums(P,na.rm=TRUE)),,drop=FALSE]
    head(P)
    
    R <- P
    if(nrow(R)>1) {
        R <- (apply(P,2,rank)/nrow(P))**4
        R <- R[order(-rowSums(R)),,drop=FALSE]
        head(R)
    }

    is.multiomics <- any(grepl("\\[gx\\]|\\[mrna\\]",rownames(R)))
    if(DEV.VERSION) is.multiomics <- (is.multiomics && input$pdx_multiomics)
    is.multiomics    
    if(is.multiomics) {
        ## EXPERIMENTAL: multi-omics weighting 
        rr <- rowMeans(R)
        dtype <- ngs$genes[rownames(R),"data_type"]
        gene  <- ngs$genes[rownames(R),"gene_name"]            
        gg <- sort(unique(gene))
        dtypes <- unique(dtype)
        dt <- "gx"
        rho <- c()
        for(dt in dtypes) {
            r1 <- rr[which(dtype == dt)]
            g1 <- gene[which(dtype == dt)]
            r1 <- r1[match(gg,g1)]
            rho <- cbind(rho, r1)
        }
        rownames(rho) <- gg
        colnames(rho) <- dtypes
        rho[is.na(rho)] <- 0
        rho <- rho[order(-rowSums(rho,na.rm=TRUE)),]
        dim(rho)            
        barplot(t(head(rho,50)), las=3)
        jj <- match(gene, rownames(rho))
        rname <- rownames(R)
        R <- rho[jj,]
        rownames(R) <- rname

    }
    
    dbg("calcVariableImportance:: 5: drawing tree\n")
    progress$inc(3/10, detail = "drawing tree")
    
    ##------------------------------
    ## create partition tree
    ##------------------------------
    require(rpart)
    R <- R[order(-rowSums(R)),,drop=FALSE]
    sel <- head(rownames(R),100)
    sel <- head(rownames(R),NFEATURES)  ## top50 features
    sel <- intersect(sel,rownames(X))
    tx <- t(X[sel,,drop=FALSE])
    dim(tx)
    
    ## formula wants clean names, so save original names
    colnames(tx) <- gsub("[: +-.,]","_",colnames(tx))
    colnames(tx) <- gsub("[')(]","",colnames(tx))
    colnames(tx) <- gsub("\\[|\\]","",colnames(tx))
    orig.names <- sel
    names(orig.names) <- colnames(tx)
    jj <- names(y)
    ##ny <- length(unique(y))
    ##if(length(jj) < ny*20) jj <- c(jj,jj,jj)
    
    if(do.survival) {
        require(survival)
        time <- abs(y)
        status <- (y>0)    ## dead if positive time
        df <- data.frame( time=time+0.001, status=status, tx)
        ##df <- df[jj,]
        rf <- rpart( Surv(time,status) ~ ., data=df)
    } else {
        df <- data.frame( y=y, tx)
        ##df <- df[jj,]
        rf <- rpart( y ~ ., data=df)
    }
    table(rf$where)
    rf$cptable
    rf$orig.names <- orig.names
    
    rf.nsplit <- rf$cptable[,"nsplit"]
    max(rf.nsplit)
    length(unique(y)) 
    if(grepl("survival",ct)) {
        MAXSPLIT = 4  ## maximum five groups....
    } else {
        MAXSPLIT = 1.5*length(unique(y))  ## maximum N+1 groups
    }
    if( max(rf.nsplit) > MAXSPLIT) {
        cp.idx <- max(which(rf.nsplit <= MAXSPLIT))
        cp0 <- rf$cptable[cp.idx,"CP"]
        ##rf <- prune(rf, cp=0.05)
        rf <- prune(rf, cp=cp0)
    }
    table(rf$where) 
    
    dbg("calcVariableImportance:: *** done!!! ***\n")
    progress$inc(2/10, detail = "done")

    ##y <- y[rownames(ngs$samples)]
    ##tx <- tx[names(y),]
    y <- y[rownames(tx)]
    colnames(tx) <- orig.names[colnames(tx)]
    ##res <- list(P=P, R=R, y=y, X=t(tx), rf=rf)
    res <- list(R=R, y=y, X=t(tx), rf=rf)
    return(res)
})


pdx_importance.RENDER <- reactive({
    
    res <- calcVariableImportance()
    if(is.null(res)) return(NULL)
    R <- res$R
    R <- R[order(-rowSums(R,na.rm=TRUE)),,drop=FALSE]
    R <- pmax(R,0.05)

    if(input$pdx_level=="geneset") {
        par(mfrow=c(1,2), oma=c(1,1,1,1)*0.5, mgp=c(2.2,0.8,0))
        par(mar=c(4,8,2,4))
        frame()
        R.top <- head(R,15)
        rownames(R.top) <- tolower(rownames(R.top))
        rownames(R.top) <- substring(rownames(R.top),1,60)
        barplot( t(R.top), las=1, horiz=TRUE,
                cex.names=0.95, xlab="cumulative importance" )
        klr <- grey.colors(ncol(R))
        legend("topright",
               legend = colnames(R), fill=klr,
               cex=0.75, y.intersp=0.8, inset=c(-0.25,0), xpd=NA )

    } else {
        par(mfrow=c(1,1), oma=c(1,1,1,1)*0.2)
        par(mar=c(7,4,0,4))
        R.top <- head(R,40)
        barplot( t(R.top), las=3, horiz=FALSE,
                cex.names=0.9, ylab="cumulative importance" )
        klr <- grey.colors(ncol(R))
        legend("topright",
               legend=rev(colnames(R)), fill=rev(klr),
               cex=0.8, y.intersp=0.8 )
    }
    ## title("Variable importance", cex.main=1.2, line=1.2, adj=0.3)    
})


pdx_heatmap.RENDER <- reactive({

    ngs <- inputData()
    ##if(is.null(ngs)) return(NULL)
    req(ngs)
    
    res <- calcVariableImportance()
    if(is.null(res)) return(NULL)

    cat("<predict:pdx_heatmap> called\n")
    
    gg <- NULL
    if(input$pdx_level=="geneset") {
        gg <- rownames(res$X)
        gg <- intersect(gg, rownames(ngs$gsetX))
        X <- ngs$gsetX[gg,]
        rownames(X) <- tolower(rownames(X))
    } else {
        gg <- rownames(res$X)
        gg <- intersect(gg, rownames(ngs$X))
        X <- ngs$X[gg,]
    }

    dbg("<predict:pdx_heatmap> dim(X)=",dim(X),"\n")
    X <- head(X[order(-apply(X,1,sd)),],40)  ## top50

    splitx <- NULL
    ct <- input$pdx_predicted
    do.survival <- grepl("survival",ct,ignore.case=TRUE)
    dbg("calcVariableImportance","do.survival = ",do.survival,"\n")

    splitx <- ngs$Y[colnames(X),ct]
    dbg("<predict:pdx_heatmap> 1:table(splitx)=",table(splitx),"\n")
    if(!is.categorical(splitx) || do.survival) {
        splitx <- NULL
    }
    dbg("<predict:pdx_heatmap> 2:table(splitx)=",table(splitx),"\n")
        
    rownames(X) <- substring(rownames(X),1,40)    
    annot <- ngs$Y[colnames(X),]
    dbg("<predict:pdx_heatmap> dim(annot)=",dim(annot),"\n")
    
    gx.splitmap( X, split=NULL, splitx=splitx,
                show_colnames = FALSE,  ## save space, no sample names
                show_legend = ifelse(is.null(splitx),TRUE,FALSE),
                col.annot=annot, annot.ht=2.5, show_rownames=50,
                lab.len=50, cexRow=0.85, mar=c(2,8))

    dbg("<predict:pdx_heatmap> done!\n")
    
})    

pdx_decisiontree.RENDER <- reactive({

    dbg("<pdx_decisiontree.RENDER> reacted")
    
    res <- calcVariableImportance()
    if(is.null(res)) return(NULL)

    require(rpart)
    require(rpart.plot)
    par(mfrow=c(1,1), mar=c(1,0,2,0))

    is.surv <- grepl("Surv",res$rf$call)[2]
    is.surv
    if(is.surv) {
        require(survival)
        require("partykit")
        (rf <- as.party(res$rf))
        ##(rf <- as.party(prune(res$rf, cp=0.05)))
        plot(rf)
        ##title("Survival tree",cex=1.2,line=0.9,adj=0.35)
        ##table(res$rf$where)
    } else {
        ##visNetworkOutput("pdx_decisiontreeClass")
        ##plotOutput("pdx_decisiontreeClass")
        ##require(visNetwork)
        if(1) {
            rpart.plot(res$rf)
            title("Classification tree",cex=1.2,line=3,adj=0.35)
        } else {
            visTree(res$rf, main="Classification tree",width="100%",legend=FALSE) %>%
                visInteraction(tooltipStyle='position:fixed;visibility:hidden;padding:5px;white-space:nowrap;font-family:helvetica;font-size:10px;background-color:lightgrey;')
        }
    } 

})
    

pdx_boxplots.RENDER <- reactive({
    res <- calcVariableImportance()
    if(is.null(res)) return(NULL)

    dbg("pdx_boxplots:: called\n")    
    vars <- setdiff(res$rf$frame$var,"<leaf>")
    vars <- res$rf$orig.names[vars]
    if(length(vars)==0) return(NULL)
    
    ##vars0 <- setdiff(vars,rownames(res$X))
    ##dbg("pdx_boxplots:: vars0=",vars0,"\n")
    if(input$pdx_level=="geneset") {
        ##xvars <- res$rf$orig.names[vars]
        vars <- intersect(vars,rownames(res$X))
    } else {
        vars <- intersect(vars,rownames(res$X))
    }

    ## add some other variables
    if(1 && length(vars)<8) {
        jj <- order(-rowSums(res$R,na.rm=TRUE))
        other.vars <- rownames(res$R)[jj]
        vars <- head(unique(c(vars,other.vars)),8)
    }
    
    y <- res$y
    is.surv <- grepl("Surv",res$rf$call)[2]
    is.surv
    if(is.surv) 
    {
        y <- paste0("N",res$rf$where)
        ##y <- as.integer(factor(y, levels=sort(unique(y))))
        names(y) <- colnames(res$X)
        table(y)
    }

    ny <- length(unique(y))    
    par(mfrow=c(2,4), mar=c(3.5,3,2,0.5),
        mgp=c(2,0.8,0), oma=c(0.5,1,1,1)*0)
    if(length(vars)>8) par(mfrow=c(3,4), mar=c(2.8,3,2,0.3),)
    i=1
    for(i in 1:min(12,length(vars))) {
        ##g <- rownames(res$R)[i]
        g <- vars[i]
        gx <- res$X[g,]
        boxplot( gx ~ y, col="grey85", ylim = range(gx),
                ylab="expression", cex.axis=0.001)
        axis(2, cex.axis=0.9)
        ##axis(1, at=1:ny, labels=levels(factor(y)), cex.axis=0.6)
        cex1 <- ifelse( ny >= 8, 0.65, 0.8)
        title(g, cex.main=ifelse(nchar(g)>20,0.85,1))
        nchar(y)
        too.big = (max(nchar(y))>=6 && ny==2) ||
            (max(nchar(y))>=4 && ny>2 || ny>=6 )
        if(too.big) {
            dy <- min(gx) - 0.16*diff(range(gx))
            nn <- max(nchar(y))
            text(1:ny, dy, levels(factor(y)), xpd=NA,
                 cex=cex1, srt=30, adj=1)
        } else {
            mtext(levels(factor(y)), side=1, line=0.7,
                  cex=cex1*0.85, las=1, at=1:ny)
        }

    }

})


pdx_importance.opts = tagList()
pdx_importance_module <- plotModule(
    "pdx_importance", pdx_importance.RENDER,
    title = "Variable importance",
    info.text = "A variable importance score for each feature is calculated using multiple machine learning algorithms, including LASSO, elastic nets, random forests, and extreme gradient boosting. By combining several methods, the platform aims to select the best possible biomarkers. The top features are plotted according to cumulative ranking by the algorithms.",
    ##options = pdx_importance.opts,
    label="a",
    pdf.width=10, pdf.height=5, ## res=80
)
output <- attachModule(output, pdx_importance_module)


pdx_heatmap.opts = tagList()
pdx_heatmap_module <- plotModule(
    "pdx_heatmap", pdx_heatmap.RENDER,
    title = "Heatmap", label="b",
    info.text = "The heatmap shows the expression of genes of the top most important features.",
    ##options = pdx_heatmap.opts,
    pdf.width=10, pdf.height=10, res=72
)
output <- attachModule(output, pdx_heatmap_module)

pdx_decisiontree.opts = tagList()
pdx_decisiontree_module <- plotModule(
    "pdx_decisiontree", pdx_decisiontree.RENDER,
    title = "Decision tree", label="c",
    info.text = "The decision tree shows a tree solution for classification based on the top most important features.",
    ##options = pdx_decisiontree.opts,
    pdf.width=10, pdf.height=6, res=72
)
output <- attachModule(output, pdx_decisiontree_module)

pdx_boxplots.opts = tagList()
pdx_boxplots_module <- plotModule(
    "pdx_boxplots", pdx_boxplots.RENDER,
    title = "Biomarker expression", label="d",
    info.text = "These boxplots shows the expression of genes/samples of the identified features.",
    ##options = pdx_boxplots.opts,
    pdf.width=10, pdf.height=8, res=90
)
output <- attachModule(output, pdx_boxplots_module)


fillRow(flex = c(1,0.1,1), 
        fillCol(
            flex = c(0.7,1),
            moduleWidget(pdx_importance_module),
            moduleWidget(pdx_heatmap_module)
        ),
        br(), ## spacer
        fillCol(
            flex = c(1,0.9), 
            moduleWidget(pdx_decisiontree_module),
            moduleWidget(pdx_boxplots_module)
        ))

```

> **Biomarker selection**. The expression of certain genes may be used as *markers* to predict a certain phenotype such as response to a therapy. Finding such *biomarkers* are of high importance in clinical applications. **(a)** An importance score for each feature is calculated using multiple machine learning algorithms, including LASSO, elastic nets, random forests, and extreme gradient boosting. The top features are plotted  according to cumulative ranking by the algorithms. **(b)** The heatmap shows the expression distribution for the top most important features. **(c)** The decision tree shows (one) tree solution for classification based on the top most important features. **(d)** Boxplots show the expression of biomarker genes across the groups.


`r if(!DEV.VERSION) {"<!---------------------"}`

### Multi-level (dev)
    
```{r message=FALSE, warning=FALSE, fig.width=4}

mix_computeDiablo <- reactive({
    ngs <- inputData()
    ##if(is.null(ngs)) return(NULL)
    req(ngs)
    
    cat("<mix_computeDiablo> called\n")

    comparison=1
    comparison="ABC_vs_GCB"
    comparison="treatment"
    comparison="cell.type"
    comparison <- input$pdx_predicted
    if(is.null(comparison)) return(NULL)
    if(comparison=="") return(NULL)
    cat("<mix_computeDiablo> comparison=",comparison,"\n")
    
    if(comparison %in% colnames(ngs$samples)) {
        Y <- ngs$samples[,comparison]
        Y[which(Y %in% c(NA,"","NA"))] <- NA
    } else {
        return(NULL)
    }
        
    ct1 <- "COMPARTMENTS:Cell_surface"
    ct2 <- "MSIGDB:PROTEIN_KINASES"
    ct3 <- "CUSTOM:Transcription factors"
    ct1 <- input$mix_level1
    ct2 <- input$mix_level2
    ct3 <- input$mix_level3
    if(is.null(ct1) || is.null(ct2) || is.null(ct3)) return(NULL)
    
    require(mixOmics)
    cs <- GSETS[[ct1]]
    pk <- GSETS[[ct2]]
    tf <- GSETS[[ct3]]
    gg <- rownames(ngs$X)
    cs.X <- ngs$X[intersect(cs, gg),]
    pk.X <- ngs$X[intersect(pk, gg),]
    tf.X <- ngs$X[intersect(tf, gg),]
    
    data = list(cs.X, pk.X, tf.X)
    names(data) <- c("CS","PK","TF")
    names(data) <- sub(".*:","",c(ct1, ct2, ct3))
    names(data) <- substring(toupper(names(data)),1,30)
    length(data)

    posonly=FALSE
    posonly <- input$pdx_posonly
    res <- pgx.makeTriSystemGraph(
        data, Y, nfeat=25, numedge=100, posonly=posonly)    
    return(res)
})


mix_plotLoadings.RENDER <- reactive({    

    cat("<predict::mix_plotLoadings> called\n")    
    ngs <- inputData()
    ##if(is.null(ngs)) return(NULL)
    req(ngs)
    
    res <- mix_computeDiablo()
    ##plotLoadings(res, comp = 2, contrib = 'max', method = 'median',
    ## size.name = 1, size.title = rel(0.7))	   
    if(is.null(res)) return(NULL)

    showloops=FALSE
    showloops <- input$mix_showloops
    mixPlotLoadings(res, showloops=showloops) 
    
})


output$mix_plotDiablo <- renderPlot({
    res <- mix_computeDiablo()
    if(is.null(res)) return(NULL)
    
    ny <- length(unique(res$Y))
    klrpal <- c("blue2","orange2")
    klrpal <- brewer.pal(n=3, name = "RdBu")[c(3,1)]
    if(ny>2) klrpal <- rep(brewer.pal(n=8, "Set2"),10)[1:ny]
    klrpal
    klr1 <- klrpal[factor(res$Y)]
    klr1[is.na(klr1)] <- "grey85"
    k=1
    if(0) {
        plotDiablo(res, ncomp = k)
    } else {             
        ## plot samples
        par(oma=c(1,1,1,1)*0, mar=c(0,0,0,0))
        vx1 <- sapply(res$variates,function(x) x[,k])
        colnames(vx1) <- substring(colnames(vx1),1,4)
        pairs(vx1,cex=1, cex.labels=1.1, pch=19, col=klr1)
    }    
})


output$mix_plotVariates <- renderPlot({
    res <- mix_computeDiablo()
    if(is.null(res)) return(NULL)
     
    ny <- length(unique(res$Y))
    klrpal <- c("blue2","orange2")
    klrpal <- brewer.pal(n=3, name = "RdBu")[c(3,1)]
    if(ny>2) klrpal <- rep(brewer.pal(n=8, "Set2"),10)[1:ny]
    klrpal
    klr1 <- klrpal[factor(res$Y)]
    klr1[is.na(klr1)] <- "grey85"
    
    ## plot samples
    par(mfrow=c(2,2), oma=c(1,1,1,1)*0, mar=c(3,3.5,2.5,0))
    par(mgp=c(1.8,0.8,0))
    for(k in 1:3) {
        vx1 <- res$variates[[k]][,1:2]
        plot(vx1, cex=1.5, pch=19, col=klr1,
             xlab="component.1", ylab="component.2")
        title(names(res$variates)[k],cex.main=1, line=0.5)
    }

    ## combining all levels
    V <- do.call(rbind, res$loadings[1:3])
    X <- do.call(cbind, res$X)[,rownames(V)]
    R <- X %*% V
    plot(R[,1:2], cex=1.5, pch=19, col=klr1,
         xlab="component.1", ylab="component.2")
    title("COMBINED",cex.main=1, line=0.5)
    
})

mix_plotSNF.RENDER <- reactive({
    res <- mix_computeDiablo()
    if(is.null(res)) return(NULL)
    
    cat("<predict::mix_plotSNF> called\n")

    ##if(!require(SNFtool)) install.packages("SNFtool")
    require(SNFtool)
    M.list <- list()
    i=1
    for(i in 1:3) {
        ##dist1 <- as.matrix(dist(res$X[[i]]))
        dist1 <- as.matrix(as.dist(1 - cor(t(res$X[[i]]))))
        k1 = floor(min(20,0.5*ncol(dist1)))
        M1 <- affinityMatrix(dist1, K=k1, sigma=0.5)
        M.list[[i]] <- M1
    }
    names(M.list) <- names(res$X)[1:3]
    
    prepMatrix <- function(M) {
        ## matrix preprocessing as in displayClustersWithHeatmap()
        normalize <- function(X) X/rowSums(X)
        diag(M) <- median(as.vector(M))
        M <- normalize(M)
        M <- M + t(M)
        return(M)
    }
    Wall <- lapply(M.list, prepMatrix)

    if(0) {
        ##dist1 <- as.matrix(dist(Data1))
        W1 <- affinityMatrix(dist1, K = 20, sigma = 0.5)
        ind <- spectralClustering(W1, K=3)
        displayClustersWithHeatmap(W1, ind)
        
        W1 <- Wall[[1]]
        ind <- order(spectralClustering(W1, K=3))
        heatmap(W1[ind,ind], scale="none", Rowv=NA, Colv=NA)
        heatmap(W1, scale="none")
       
    }

    k1 = floor(min(20,0.5*ncol(Wall[[1]])))
    Wcomb = SNF(Wall, K=k1, t=20)
    Wcomb = prepMatrix(Wcomb)
    estK = estimateNumberOfClustersGivenGraph(Wcomb, NUMC=2:8)
    estK
    ##K=5
    K=-1
    ##K=median(unlist(estK))
    K    

    ind <- hclust(as.dist(1 - cor(Wcomb)))$order
    if(K>0) ind <- order(spectralClustering(Wcomb, K=K))
    
    par(mfrow=c(2,2), mar=c(0.5,1,2,0.5))
    i=1
    for(i in 1:3) {
        ##ind <- hclust(as.dist(1 - cor(Wall[[i]])))$order
        ##if(K>0) ind <- order(spectralClustering(Wall[[i]], K=K))
        ##heatmap(Wall[[i]][ind,ind], scale="none", Rowv=NA, Colv=NA)
        image(Wall[[i]][ind,ind], xaxt='n', yaxt='n')
        title(names(Wall)[i],cex.main=1.1,line=0.5)
    }

    ##displayClustersWithHeatmap(W, spectralClustering(W, K=3))
    ##heatmap(prepMatrix(W), scale="none")
    image(Wcomb[ind,ind], xaxt='n', yaxt='n')
    title("SNF",cex.main=1.1,line=0.5)
    
})

output$mix_circosPlot <- renderPlot({
    res <- mix_computeDiablo()
    if(is.null(res)) return(NULL)
    
    circosPlot(res, cutoff = 0.95, line = TRUE, 
               color.blocks= c('darkorchid', 'brown1', 'lightgreen'),
               color.cor = c("grey30","red3"), size.labels = 1.5)

})
    

mix_hivePlot.RENDER <- reactive({

    ngs <- inputData()
    ##if(is.null(ngs)) return(NULL)
    req(ngs)

    ct="group"
    ct <- input$pdx_predicted
    if(is.null(ct)) return(NULL)

    ##-------------------------------------------------------------
    ## Get the DIABLO results
    ##-------------------------------------------------------------
    res <- mix_computeDiablo()
    if(is.null(res)) return(NULL)

    showloops = FALSE
    showloops = input$mix_showloops
    mixHivePlot(res=res, ngs=ngs, ct=ct, showloops=showloops)

})


mix_cimDiablo.RENDER <- reactive({
    ngs <- inputData()
    ## if(is.null(ngs)) return(NULL)
    req(ngs)

    res <- mix_computeDiablo()
    if(is.null(res)) return(NULL)

    NFEAT=20
    top.features <- lapply(res$loadings[1:3], function(x)
        head(names(sort(-abs(rowMeans(x**2)))),NFEAT)
        )
    X <- do.call(cbind, res$X)
    X <- X[,unlist(top.features)]
    dim(X)
    X <- X[,head(order(-apply(X,2,sd)),60)]
    idx <- as.integer(sub(":.*","",colnames(X)))
    ptype <- names(res$variates)[idx]
    col.annot <- ngs$Y[rownames(X),]
    colnames(X) <- sub(".*[:]","",colnames(X))
    
    gx.splitmap( t(X), split=ptype, splitx=NULL,
                softmax=1, scale="row", 
                col.annot=col.annot, cexRow=0.9, 
                dist.method="pearson",                
                col.dist.method="pearson")
})


mix_plotLoadings.opts = tagList()
mix_plotLoadings_module <- plotModule(
    "mix_plotLoadings", mix_plotLoadings.RENDER,
    title = "Multi-level variable importance", label="a",
    info.text = "A multi-level variable importance score for each feature is calculated using the DIABLO framework from the MixOmics package.",
    ## options = mix_plotLoadings.opts,
    pdf.width=10, pdf.height=10, res=80
)
output <- attachModule(output, mix_plotLoadings_module)

mix_hivePlot.opts = tagList()
mix_hivePlot_module <- plotModule(
    "mix_hivePlot", mix_hivePlot.RENDER,
    title = "Hive plot ", label="b",
    info.text = "The hive plot is a novel visualization method for drawing networks (see http://www.hiveplot.com). The purpose of the hive plot is to visually exploring the network structure. Here, size of the nodes represent the variable importance of the genes and are radially positioned based on the fold-change or F-stats. Edges are drawn as curved links and their thickness represent edge importance. By turning on 'detect loops' in the options, one can highlight possible autocrine signaling loops.",
    ##options = mix_hivePlot.opts,
    pdf.width=10, pdf.height=10, res=80
)
output <- attachModule(output, mix_hivePlot_module)


mix_plotSNF.opts = tagList()
mix_plotSNF_module <- plotModule(
    "mix_plotSNF", mix_plotSNF.RENDER,
    title = "SNF clustering", label="c",
    info.text = "SNF is a novel computational method for genomic data integration. SNF constructs patient similarity networks for each of the data types and in a second step integrates iteratively these networks until it converges to a final fused network. SNF is able to deal with various data types and since the integration happens in the sample space one can combine various data type without worrying for different scales or noise among the different data. For more information see Wang et al., Nature methods, 2014.",
    ##options = mix_plotSNF.opts,
    pdf.width=10, pdf.height=10, res=80
)
output <- attachModule(output, mix_plotSNF_module)


mix_cimDiablo.opts = tagList()
mix_cimDiablo_module <- plotModule(
    "mix_cimDiablo", mix_cimDiablo.RENDER,
    title = "Expression heatmap", label="d",
    info.text = "The heatmap shows the expression values of the most important features.",
    ##options = mix_cimDiablo.opts,
    pdf.width=10, pdf.height=10, res=72
)
output <- attachModule(output, mix_cimDiablo_module)

plotFill <- function(module, id) {
    fillCol(flex = c(NA,1),
            module$buttons,
            plotOutput(id)
            )
}

mix_options <- tagList(
    tipify(checkboxInput('mix_showloops','detect signaling loops',FALSE),
           "Check to enable signaling loops detection. Signaling loops are detected by finding highly correlated paths that connect all three levels and also show a feedback. These loops may suggest self-sustaining autocrine loops."),
    tipify(checkboxInput('pdx_posonly','positive correlation only',FALSE),
           "Restrict to only positive correlation.")
)
mix_buttons <- plotModuleButtons(
    "mix_level", text="Multi-level variable selection. This module is based on the DIABLO framework of the MixOmics R/package.",
    no.download=TRUE, options=mix_options)

fillCol(flex = c(NA,1),
        inputPanel(

            tipify( selectInput("mix_level1",NULL,choices=NULL), placement="right",
                   "Select first level of biomarkers. These are often signaling genes on the cell surface."),
            
            tipify( selectInput("mix_level2",NULL,choices=NULL), placement="right",
                   "Select second level of biomarkers. These are often genes relaying the signals to the effector genes."),

            tipify( selectInput("mix_level3",NULL,choices=NULL), placement="right",
                   "Select third level of biomarkers. These are often effector genes such astranscription factors."),
            mix_buttons,            
            cellArgs=list(width='350px')
        ),                
        fillRow(flex = c(1.8,0.1,1),
                fillCol(flex = c(0.9,0.06,1),
                        fillCol(flex = c(NA,1),
                                mix_plotLoadings_module$buttons,
                                plotOutput('mix_plotLoadings')  %>% withSpinner(proxy.height="360px")
                                ),
                        br(),
                        fillRow(flex = c(1.2,1),
                                ##plotOutput('mix_plotDiablo'),
                                fillCol(flex = c(NA,1),
                                        mix_hivePlot_module$buttons,
                                        plotOutput('mix_hivePlot') %>% withSpinner(proxy.height="380px")
                                        ),
                                ##plotOutput('mix_plotVariates')
                                fillCol(flex = c(NA,1),
                                        mix_plotSNF_module$buttons,
                                        plotOutput('mix_plotSNF') %>% withSpinner(proxy.height="380px")
                                        )
                                ##plotOutput('mix_circosPlot')
                                )
                        ),
                br(),
                fillCol(flex = c(NA,1),
                        mix_cimDiablo_module$buttons,
                        plotOutput('mix_cimDiablo')  %>% withSpinner(proxy.height="750px")
                        )
                )
        )

observe({
    ngs <- inputData()
    ##if(is.null(ngs)) return(NULL)
    req(ngs)
    ct1 <- c("COMPARTMENTS:Cell_surface", "COMPARTMENTS:Membrane_part",
             "FAMILY:Endogenous ligands (HGNC)",
             "C2:REACTOME_CHEMOKIoNE_RECEPTORS_BIND_CHEMOKINES")
    ct2 <- c("COMPARTMENTS:Cytoplasm", "MSIGDB:PROTEIN_KINASES",
             "C5:GO_KINASE_ACTIVITY","FAMILY:Kinases (KEA)",
             "FAMILY:Micro RNA")
    ct3 <- c("COMPARTMENTS:Nuclear_part","CUSTOM:Transcription factors",
             "FAMILY:Nuclear receptors","MSIGDB:TRANSCRIPTION_FACTORS")
    gs.size <- function(gs) {length(intersect(gs,rownames(ngs$X)))}
    ct1 <- ct1[which(sapply(GSETS[ct1],gs.size)>=3)]
    ct2 <- ct2[which(sapply(GSETS[ct2],gs.size)>=3)]
    ct3 <- ct3[which(sapply(GSETS[ct3],gs.size)>=3)]
    updateSelectInput(session, "mix_level1", choices=ct1)
    updateSelectInput(session, "mix_level2", choices=ct2)
    updateSelectInput(session, "mix_level3", choices=ct3)
})

```

> **Multi-level feature selection.** **(a)** A multi-level importance score for each feature is calculated using the DIABLO framework from  the [MixOmics](http://mixomics.org/) package. The DIABLO method generalises PLS for multiple datasets, here applied to compartment gene sets. **(b)** The Hive plot visualized the connectivity of the network. **(c)** SNF constructs similarity networks for each of the levels and integrates these networks to a final fused network. **(d)** The heatmap shows the expression values of the most important features in each of the multiple levels.


`r if(!DEV.VERSION) {"--------------------->"}`
